### 2024-04-22 manuscript/SI_temp.Rmd
files pre-publication for movement to christeLab repo

---
title: "Untitled"
output:
  word_document: default
  html_document: default
date: "2024-04-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

Install:

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("hemstrow/snpR")
```

load
```{r}
library(snpR)
```

download

```{r}
# download the metadata
meta <- read.table(url("https://raw.githubusercontent.com/hemstrow/seg_sites_rarefaction/main/data/example_metadata.txt"), 
                   header = TRUE)

# download the vcf
download.file("https://raw.githubusercontent.com/hemstrow/seg_sites_rarefaction/main/data/example_vcf.vcf", 
              destfile = "example_vcf.vcf")
```

read in
```{r}
monarchs <- read_vcf("example_vcf.vcf", sample.meta = meta)
```

types
```{r}
summarize_facets(monarchs, "pop")
monarchs <- calc_seg_sites(monarchs, "pop", g = 0)
get.snpR.stats(monarchs, "pop", "seg_sites")$weighted.means

monarchs <- calc_seg_sites(monarchs, "pop", g = -1)
get.snpR.stats(monarchs, "pop", "seg_sites")$weighted.means

monarchs <- calc_seg_sites(monarchs, "pop", g = 10)
get.snpR.stats(monarchs, "pop", "seg_sites")$weighted.means

monarchs <- calc_seg_sites(monarchs, "pop", g = 0)
head(get.snpR.stats(monarchs, "pop", "seg_sites")$single, 10)
```