### 2024-04-22 manuscript/ms.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setsansfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage[left]{lineno}
\linenumbers
\usepackage{setspace}\doublespacing
\usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
\usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
\usepackage[round]{natbib}
\usepackage{amsmath}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={manuscript},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{manuscript}
\author{}
\date{\vspace{-2.5em}31 October 2023}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{probability-of-observing-segregating-loci-via-rarefaction}{%
\subsection{Probability of observing segregating loci via
rarefaction}\label{probability-of-observing-segregating-loci-via-rarefaction}}

Rarefaction can be used to estimate the probability of observing a
segregating site at a specific locus using much the same framework used
to calculate allelic richness. In brief, allelic richness (or the
expected number of distinct alleles expected at a given locus under a
common sample size \(g\) across populations) can be estimated for a
given population \(j\) by summing the probability of observing each
\(i\) of \(m\) unique alleles using the counts of those alleles in the
population \(N_{ij}\) and the total sample size in that population
\(N_{j}\). This is done by comparing the number of possible ways to draw
\(g\) gene copies without sampling allele \(i\)
(\(\binom{N_{j} - N_{ij}}{g}\)) to the total number of possible
combinations of gene copies that can be drawn (\(\binom{N_{j}}{g}\));
the inverse of this
(\(1 - \frac{\binom{N_{j} - N_{ij}}{g}}{\binom{N_{j}}{g}}\)) is
therefore the probability of observing allele \(i\) in population
\(N_{j}\), and the sum of this value across all \(m\) alleles gives the
expected number of alleles observed at a locus in population \(j\),
\(\alpha_{g}^{j}\) (\textbf{hurlbertNonconceptSpeciesDiversity1971?};
\textbf{kalinowskiCountingAllelesRarefaction2004?}):

\begin{equation}
\alpha_{g}^{j} = \sum_{i = 1}^{m} 1 - \frac{\binom{N_{j} - N_{ij}}{g}}{\binom{N_{j}}{g}}
\end{equation}

The expected number of segregating loci in a population for a draw of
\(g\) gene copies can be derived similarly. For a locus \(i\) to be
segregating in population \(j\), all alleles drawn across all gene
copies must be identical. If alleles are independent at each locus (the
locus is at Hardy-Weinburg Equlibrium, HWE) and \(N\) is infinite, the
probability (\(P({S_{j}})\)) of observing a segregating site at a locus
is the inverse of the probability of drawing only one allele in \(g\)
draws with replacement: \[P({S_{j}}) = 1 - \sum_{i = 1}^{m} f_{ij}^{g}\]

where \(f_{ij}\) is the allele frequency of allele \(i\) in population
\(j\). However, in finite samples draws are conducted with replacement,
and so binomial coefficients must instead be used to determine the
probability of drawing only a specific allele:
\[P({S_{j}}) = 1 - \sum_{i = 1}^{m} \frac{\binom{N_{ij}}{g}}{\binom{N_{j}}{g}}\]

However, HWE is often not a desirable assumption to make. Even if
filtering is employed to remove loci which do not conform to HWE, the
degree of conformity, and thus the degree of statistical bias in
estimating \(P({S_{j}})\), typically varies somewhat between
populations. For example, in a sample of 100 genotypes with a minor
allele frequency of 0.05, only five minor alleles are expected and two
out of three possible combinations of minor homozygotes and
heterozygotes that produce that frequency will not deviate from HWE at
\(\alpha = 0.05\) according to an exact test (Wigginton et al., 2005).
However, re-sampling these to, say, ten genoytpes should will produce
quite different \(P(S_{j})\) (roughly 0.7, 0.6, and 0.5, for purely
heterozygotes, one homozygote and three heterozygotes, and two
homozygotes and one heterozygote, respectively) as we will see below.

To remedy these problems, I propose the following estimator of
\(P({S_{j}})\):
\[P({S_{j}}) = 1 - \sum_{k = 1}^{h}\frac{\binom{n_{kj}}{\gamma}}{\binom{n_{j}}{\gamma}}\]
where the \(P({S_{j}})\) is given by the probability of exclusively
drawing any \(k\) of \(h\) possible homozygote genotypes in population
\(j\) given \(\gamma\) independent sampled \emph{genotypes} (not
\emph{gene copies}) from the pool of observed genotypes. Here,
\(n_{kj}\) is the number of observed homozygote genotypes of type \(k\)
in population \(j\) and \(n_{j}\) is the total number of observed
genotypes of all types. Note that \(n_{j}\) and \(\gamma\) will be half
the value of their equivalents \(N_{j}\) and \(g\) for diploid species,
one third for triploids, and so on.

Interestingly, this method, like the richness method and related private
allele rarefaction approaches can smoothly account for varying amounts
of missing data at specific loci in different populations by varying
\(g\) or \(\gamma\) across loci. Both can be set to one less than the
smallest observed \(N_{j}\) or \(n_{j}\), the highest values at which
rarefaction can be applied within a population, across all populations
after accounting for missing data, and can thus vary across loci without
bias. Setting either value to \(N_{j}\) or \(n_{j}\) will instead return
the observed allele diversity or segregating site status, respectively.

This is particularly useful given that \(E(N_{S})\) or the expected
total number of segregating sites, is often of specific interest as a
measure of genetic diversity when comparing populations. Given that the
expected number of segregating sites at a specific locus \(q\) in
population \(j\), \(E(N_{S_{jq}})\), is equal to \(P({S_{jq}})\),
\(E(N_{S})\) can be calculated by summing \(P({S_{jq}})\) across all
\(Q\) loci: \[E(N_{S}) = \sum_{q = 1}^{Q} P({S_{jq}})\] with \(\gamma\)
set accordingly for each locus. In this case,
\(0 \le E(N_{S_{jq}}) \le 1\) for all loci (and thus
\(0 \le E(N_{S}) \le Q\)).

Usefully, under this framework each locus represents a single Bernoulli
trial in which it can be observed to be segregating or not with
probability \(P({S_{jq}})\). As such, the variance of \(P({S_{jq}})\)
for each locus is given by
\[\sigma_{P(S_{jq})}^{2} = P(S_{jq}) \times (1 - P(S_{jq}))\] and, if
each locus is independent, the variance of \(E(N_{S})\) is equal to the
sum of \(\sigma_{P(S_{jq})}^{2}\) across all loci:
\[\sigma_{E(n_{S})}^{2} = \sum_{q = 1}^{Q}\sigma_{P(S_{jq})}^{2}\]

Confidence and prediction intervals can then be derived using standard
approaches for the sum of random, independent Bernoulli trials. When
\(Q\) is large, for example, the distribution of \(E(N_{S})\) should be
approach normal and confidence and prediction intervals can be derived
using standard normal approximation using the equations
\[CI_{N_{S}} = E(N_{S}) \pm Z\sqrt{\frac{\sigma_{E(n_{S})}^{2}}{Q}}\]
and \[PI_{N_{S}} = E(N_{S}) \pm Z\sqrt{\sigma_{E(n_{S})}^{2}(1+(1/Q))}\]
where \(Z\) is given by the normal quantile function
\(Z = Q_{X}(1-\alpha)\) with \(\mu = 0\) and \(\sigma = 1\) for a
desired confidence level \(\alpha\).

\hypertarget{emperical-validation}{%
\subsection{Emperical Validation}\label{emperical-validation}}

To validate equations 4 and 5, I simulated genotypic data for two
populations with sizes 100 and 1000, each with 100 bi-allelic loci with
minor allele frequencies spaced equally between 0.01 and 0.1. I added
missing data to each population assigning each locus a missing data rate
\(R_{mq}\) from a uniform distribution such that
\(R_{mq} \sim U(0,.3)\), ensuring that overall allele frequencies in
each population were maintained. I then used the methods described above
to estimate \(P({S_{jq}})\) and \(E(N_{S})\) and their variances given
\(\gamma = 30\). For comparison, I also conducted 10,000 random draws of
size \(\gamma\) from each loci in each population, then calculated
\(P({S_{jq}})\) and its variance for each locus empirically and
\(E(N_{S})\) by summing across all 100 loci for each set of draws. I
likewise calculated the variance of \(N_{S}\) directly across all sets
of random draws.

To compare my calculated estimates to the empirical simulations, I used
the ``exact'' method from the R package \texttt{binom}
(\textbf{rajBinomBinomialConfidence2022?}) to calculate 95\% confidence
intervals for each empirical \(P({S_{jq}})\). I also used normal
approximation to calculate 95\% confidence and prediction intervals for
\(E(N_{S})\) and \(N_{S}\) using \(\sigma_{E(n_{S})}^{2}\) and the
variance directly observed from the simulations, respectively.

\hypertarget{results}{%
\section{Results}\label{results}}

The methods described here for calculating \(P({S_{j}})\) and
\(E(N_{S})\) performed well. Individual \(P({S_{jq}})\) values for each
locus were within the confidence intervals derived from the simulated
values for 97 and 96\% of loci from the \(n = 100\) and \(n = 1000\)
populations, respectively (Figure 1). Note that variation along the
generally correlated minor allele frequency/\(P({S_{jq}})\) axes is due
to variations in genotype frequencies in the simulated data for a given
minor allele frequency. \(P({S_{jq}})\) values estimated using equation
X account for this adequately.

\(E(N_{S})\) values estimated with equation X were similarly accurate
and were within the 95\% confidence intervals produced using the
simulated \(N_{S}\) for both population sizes. Likewise, the 95\%
prediction intervals calculated using \(\sigma_{E(n_{S})}^{2}\) via
equation X contained 96.1 and 95.3\% of the simulated \(N_{S}\) values
for \(n = 100\) and \(n = 1000\), respectively.

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Wigginton2005}{}}%
Wigginton, J. E., Cutler, D. J., \& Abecasis, G. R. (2005). A {Note} on
{Exact} {Tests} of {Hardy}-{Weinberg} {Equilibrium}. \emph{The American
Journal of Human Genetics}, \emph{76}(5), 887--893.
https://doi.org/\url{https://doi.org/10.1086/429864}

\end{CSLReferences}

\end{document}