### 2024-04-22 scripts/seg_sites_rarefaction.Rmd

---
title: "Seg_sites_rarefaction"
output: html_document
date: "2023-08-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE); library(snpR); library(ggplot2); library(data.table);
library(cowplot)
```

# Define parameters
```{r}
# define parameters
seed <- 231266
threads <- 25
allele_frequencies <- seq(.001, .1, by = .001) # allele frequencies
n1 <- 5000 # p1 gene copy number
n2 <- 500 # p2 gene copy number
k <- seq(10, 100, by = 10) # numbers to rarefact to
mcmcs <- seq(1000, 10000, by = 1000) # number of mcmc draws
max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist

## derived
set.seed(seed)
nal <- length(allele_frequencies)
miss_percent <- floor(runif(nal, 0, max_missing))/100

## plotting
colors <- khroma::color("highcontrast")(3)
ci_col <- as.character(colors[3])
colors <- as.character(colors[c(2,1)])
t.sizes <- c(20, 18)
```

# Setup genotypes
```{r}
# generate genotypes
g1 <- matrix(0, nrow = nal, ncol = n1) # init
g2 <- matrix(0, nrow = nal, ncol = n2)

# add missing data
ms1 <- lapply(n1*miss_percent, function(x) sample(n1, x, FALSE))  # which are missing
ms2 <- lapply(n2*miss_percent, function(x) sample(n2, x, FALSE))

# determine minor allele placements
ma1 <- ma2 <- vector("list", length(ms1))
for(i in 1:nal){
  ma1[[i]] <- sample(n1*(1-miss_percent)[i], (n1*(1-miss_percent)*allele_frequencies)[i], FALSE)
  ma2[[i]] <- sample(n2*(1-miss_percent)[i], (n2*(1-miss_percent)*allele_frequencies)[i], FALSE)

}

# populate
## function to replace minors
populate <- function(g, missing, minors){
  if(length(missing) > 0){
    g[missing] <- NA
    g[-missing][minors] <- 1
  }
  else{
    g[minors] <- 1
  }
  return(g)
}
## do for every locus
for(i in 1:nal){ 
  g1[i,] <- populate(g1[i,], ms1[[i]], ma1[[i]])
  g2[i,] <- populate(g2[i,], ms2[[i]], ma2[[i]])
}

## combine into individuals
g1 <- g1[,1:(ncol(g1)/2)] + g1[,((ncol(g1)/2) + 1):ncol(g1)]
g2 <- g2[,1:(ncol(g2)/2)] + g2[,((ncol(g2)/2) + 1):ncol(g2)]

```

# Rarefact mathematically
```{r}
counts <- function(g){
  return(data.frame(majhom = rowSums(g == 0, na.rm = TRUE), 
                    het = rowSums(g == 1, na.rm = TRUE), 
                    minhom = rowSums(g == 2, na.rm = TRUE)))
}
rarefact <- function(counts, k){
  if(length(k) > 1){
    warning("length(k) > 1, using first element only.\n")
    k <- k[1]
  }
  return(1 - ((choose(counts[,1], k) + choose(counts[,3], k))/choose(rowSums(counts), k)))
}
varseg <- function(p) p*(1-p)

c1 <- counts(g1) # geno counts
c2 <- counts(g2)

# rarefact
pseg1 <- lapply(k, function(tk) rarefact(c1, tk))
pseg2 <- lapply(k, function(tk) rarefact(c2, tk))
names(pseg1) <- names(pseg2) <- paste0("K_", k)

vseg1 <- lapply(pseg1, varseg)
vseg2 <- lapply(pseg2, varseg)

p1 <- lapply(pseg1, function(p) sum(p)/nal)
p2 <- lapply(pseg2, function(p) sum(p)/nal)
```

# Rarefact via simulations
Note: this chunk should run on a cluster or somewhere else allowing for large numbers of threads, then be re-imported in the next chunk. `q` should be passed as an argument for the iteration, as should `outfile`.

```{r cluster_MCMCs}
# set.seed(seed + q) # reset seed for cluster

rarefact_mcmc <- function(g, k){
  if(length(k) > 1){
    warning("length(k) > 1, using first element only.\n")
    k <- k[1]
  }

  miss <- is.na(g)
  if(any(miss)){
    g <- g[-which(miss)]
  }

  samp <- sample(g, k, FALSE)
  seg <- ifelse(any(samp == 1), 1, # got hets
                ifelse(any(samp == 0) & any(samp == 2), 1, 0)) # got both homs
  return(seg)
}

# # draw for each k/mcmc count combination
# combs <- expand.grid(k, mcmcs)
# colnames(combs) <- c("k", "iters")
# 
# tmcmcs <- combs$iters[q]
# tk <- combs$k[q]
# 
# seg1 <- seg2 <- matrix(0, nal, tmcmcs)
# for(i in 1:tmcmcs){
#   for(j in 1:nal){
#     seg1[j,i] <- rarefact_mcmc(g1[j,], k = tk)
#     seg2[j,i] <- rarefact_mcmc(g2[j,], k = tk)
#   }
# }
# 
# mpseg1 <- binom::binom.confint(rowSums(seg1), rep(tmcmcs, nrow(seg1)),
#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
# mpseg2 <- binom::binom.confint(rowSums(seg2), rep(tmcmcs, nrow(seg1)),
#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
# 
# colnames(mpseg1) <- colnames(mpseg2) <- c("mean", "lower", "upper")
# 
# res <- list(seg1 = seg1, seg2 = seg2, mpseg1 = mpseg1, mpseg2 = mpseg2, k = tk, iters = tmcmcs)
# 
# saveRDS(res, paste0(outfile, "_", q, ".RDS"))

```

# import and condense simulation results
```{r}
res_files <- list.files("../data/", "mcmc.+\\.RDS", full.names = TRUE)
res_files <- lapply(res_files, readRDS)
ks <- unlist(purrr::map(res_files, "k"))
n_iters <- unlist(purrr::map(res_files, "iters"))

```

# Figure 1: Pseg for four individual runs (k = 10 and 100, N = 100, N = 1000)

```{r Figure1}
prep_f1 <- function(pseg1, pseg2, mpseg1, mpseg2, n1, n2){
  #========plot the empirical CIs around every loci vs the predicted value==============
  # plotting the empirical CIs vs the estimated mean
  pseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies),
                data.frame(pop = paste0("N = ", n2/2), p = pseg2, snp = 1:nal, af = allele_frequencies))

  oseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = mpseg1[,1], snp = 1:nal, af = allele_frequencies),
                data.frame(pop = paste0("N = ", n2/2), p = mpseg2[,1], snp = 1:nal, af = allele_frequencies))

  pseg$source <- "Mathematical"
  oseg$source <- "Simulation"

  seg <- rbind(pseg, oseg)
  cis <- rbind(cbind.data.frame(mpseg1, pop = paste0("N = ", n1/2), af = allele_frequencies),
               cbind.data.frame(mpseg2, pop = paste0("N = ", n2/2), af = allele_frequencies))
  colnames(cis) <- c("p", "lower", "upper", "pop", "af")
  seg <- merge(seg, cis[,c("lower", "upper", "af", "pop")], by = c("af", "pop"))
  seg$in_ci <- ifelse(seg$p >= seg$lower & seg$p <= seg$upper, TRUE, FALSE)

  return(seg)
}

K10_data <- which(ks == 10 & n_iters == 10000)
F1data_K10 <- prep_f1(pseg1 = pseg1$K_10, 
                      pseg2 = pseg2$K_10, 
                      mpseg1 = res_files[[K10_data]]$mpseg1,
                      mpseg2 = res_files[[K10_data]]$mpseg2,
                      n1 = n1,
                      n2 = n2)

K100_data <- which(ks == 100 & n_iters == 10000)
F1data_K100 <- prep_f1(pseg1 = pseg1$K_100, 
                       pseg2 = pseg2$K_100, 
                       mpseg1 = res_files[[K100_data]]$mpseg1,
                       mpseg2 = res_files[[K100_data]]$mpseg2,
                       n1 = n1,
                       n2 = n2)

F1data_K10$K <- "gamma * ' = 10'"
F1data_K100$K <- "gamma * ' = 100'"

F1data <- rbind(F1data_K10, F1data_K100)
F1data$pop <- paste0("'", F1data$pop, "'")

F1data$in_ci <- ifelse(F1data$in_ci, "Yes", "No")

f1 <- ggplot(F1data[F1data$source == "Mathematical",],
             aes(x = af, y = p)) +
  geom_point(aes(color = in_ci), size = 2.5) +
  geom_errorbar(aes(ymax = upper, ymin = lower)) +
  facet_grid(K~pop, labeller = label_parsed) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  scale_color_manual(values = colors) +
  guides(color = guide_legend(title = "In 95% CI")) +
  ylab(bquote(italic(hat(P)*"("*S[jq]*")"))) +
  xlab("Minor Allele Frequency") +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        strip.text = element_text(size = t.sizes[1]),
        legend.title = element_text(size = t.sizes[1]),
        legend.text = element_text(size = t.sizes[2]))

ggsave("Figure1.pdf", f1, width = 15, height = 10)
tapply(F1data[F1data$source == "Mathematical", "in_ci"], F1data[F1data$source == "Mathematical", c("K", "pop")], mean) # percent in interval
```

# Figure 3: Distribution of Nseg vs Eseg for four runs
Note that the label are flipped here: this produces figure 3.

```{r}
prep_f2 <- function(seg1, seg2, vseg1, vseg2, F1data, n1, n2){
  seg <- F1data

  # normal approximation CI and PI equation given var and N
  nap_pi <- function(alpha, vseg, n) qnorm(1 - (alpha/2))*sqrt(vseg)*sqrt(1 + (1/n))
  nap_ci <- function(alpha, vseg, n) qnorm(1 - (alpha/2))*sqrt(vseg)/sqrt(n)

  # sums across loci
  msum1 <- colSums(seg1)
  msum2 <- colSums(seg2)

  # CI/PI
  msum_ci1 <- nap_ci(.05, var(msum1), length(msum1))
  msum_ci2 <- nap_ci(.05, var(msum2), length(msum2))

  sum_pi1 <- nap_pi(0.05, sum(vseg1), nal)
  sum_pi2 <- nap_pi(0.05, sum(vseg2), nal)

  # add to table
  msum_tab <- data.frame(nseg = c(mean(msum1), mean(msum2)),
                         lower = c(mean(msum1) - msum_ci1, mean(msum2) - msum_ci2),
                         upper = c(mean(msum1) + msum_ci1, mean(msum2) + msum_ci2),
                         pop = c(paste0("n = ", n1/2), paste0("n = ", n2/2)))

  sum_tab <- tapply(seg[seg$source == "Mathematical",]$p, seg[seg$source == "Mathematical",]$pop, sum)
  sum_tab <- as.data.frame(sum_tab)
  sum_tab$pop <- tolower(rownames(sum_tab))
  colnames(sum_tab)[1] <- "nseg"
  sum_tab <- merge(sum_tab, msum_tab[,c("pop", "lower", "upper")], by = "pop")
  sum_tab$in_ci <- sum_tab$nseg <= sum_tab$upper & sum_tab$nseg >= sum_tab$lower

  # add pi
  all_mcmc_sums <- rbind(cbind.data.frame(nseg = msum1, pop = paste0("n = ", n1/2)),
                         cbind.data.frame(nseg = msum2, pop = paste0("n = ", n2/2)))
  pi_tab <- data.frame(pop = c(paste0("n = ", n1/2), paste0("n = ", n2/2)),
                       lower = c(sum_tab$nseg[sum_tab$pop == paste0("n = ", n1/2)] - sum_pi1,
                                 sum_tab$nseg[sum_tab$pop == paste0("n = ", n2/2)] - sum_pi2),
                       upper = c(sum_tab$nseg[sum_tab$pop == paste0("n = ", n1/2)] + sum_pi1,
                                 sum_tab$nseg[sum_tab$pop == paste0("n = ", n2/2)] + sum_pi2))
  all_mcmc_sums <- merge(all_mcmc_sums, pi_tab, by = "pop")
  all_mcmc_sums$in_pi <- all_mcmc_sums$nseg <=  all_mcmc_sums$upper & all_mcmc_sums$nseg >=  all_mcmc_sums$lower

  return(list(sum_tab = sum_tab, all_mcmc_sums = all_mcmc_sums, pi_tab = pi_tab))
}

K20_data <- which(ks == 20 & n_iters == 10000)
F1data_K20 <- prep_f1(pseg1 = pseg1$K_20, 
                       pseg2 = pseg2$K_20, 
                       mpseg1 = res_files[[K20_data]]$mpseg1,
                       mpseg2 = res_files[[K20_data]]$mpseg2,
                       n1 = n1,
                       n2 = n2)

K10_data <- which(ks == 10 & n_iters == 10000)
F2data_K10 <- prep_f2(seg1 = res_files[[K10_data]]$seg1, 
                      seg2 = res_files[[K10_data]]$seg2, 
                      vseg1 = vseg1$K_10,
                      vseg2 = vseg2$K_10,
                      F1data = F1data_K10,
                      n1 = n1,
                      n2 = n2)

K20_data <- which(ks == 20 & n_iters == 10000)
F2data_K20 <- prep_f2(seg1 = res_files[[K20_data]]$seg1, 
                       seg2 = res_files[[K20_data]]$seg2, 
                       vseg1 = vseg1$K_20,
                       vseg2 = vseg2$K_20,
                       F1data = F1data_K20,
                       n1 = n1,
                       n2 = n2)

F2data_K10$sum_tab$K <- "gamma * ' = 10'"
F2data_K20$sum_tab$K <- "gamma * ' = 20'"
F2data_K10$all_mcmc_sums$K <- "gamma * ' = 10'"
F2data_K20$all_mcmc_sums$K <- "gamma * ' = 20'"
F2data_K10$pi_tab$K <- "gamma * ' = 10'"
F2data_K20$pi_tab$K <- "gamma * ' = 20'"

F2data_sums <- rbind(F2data_K10$sum_tab, F2data_K20$sum_tab)
F2data_sums$pop <- gsub("n", "N", F2data_sums$pop)
F2data_sums$pop <- paste0("'", F2data_sums$pop, "'")

F2data_mcmcs <- rbind(F2data_K10$all_mcmc_sums, F2data_K20$all_mcmc_sums)
F2data_mcmcs$pop <- gsub("n", "N", F2data_mcmcs$pop)
F2data_mcmcs$pop <- paste0("'", F2data_mcmcs$pop, "'")

F2data_pi <- rbind(F2data_K10$pi_tab, F2data_K20$pi_tab)
F2data_pi$pop <- gsub("n", "N", F2data_pi$pop)
F2data_pi$pop <- paste0("'", F2data_pi$pop, "'")

F2data_mcmcs <- as.data.table(F2data_mcmcs)
levs <- unique(F2data_mcmcs[,c("pop", "K")])
F2data_densities <- vector("list", nrow(levs))

for(i in 1:nrow(levs)){
  d <- density(F2data_mcmcs[K == levs$K[i] & pop == levs$pop[i],]$nseg)
  q <- quantile(F2data_mcmcs[K == levs$K[i] & pop == levs$pop[i],]$nseg, c(0.025,0.975))
  F2data_densities[[i]] <- data.table(x = d$x, y = d$y, K = levs$K[i], pop = levs$pop[i])
  F2data_densities[[i]]$quant <- factor(findInterval(F2data_densities[[i]]$x, q))
}

F2data_densities <- rbindlist(F2data_densities)

f2 <- ggplot(F2data_densities, aes(x = x, y = y)) +
  geom_line() +
  geom_ribbon(aes(ymin = 0, ymax = y, fill = quant), show.legend = FALSE) +
  geom_vline(data = F2data_sums, aes(xintercept = nseg), color = colors[1]) +
  geom_vline(data = F2data_pi, aes(xintercept = lower), linetype = "dashed", color = colors[2]) +
  geom_vline(data = F2data_pi, aes(xintercept = upper), linetype = "dashed", color = colors[2]) +
  facet_grid(K~pop, labeller = label_parsed) +
  theme_bw() +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("#004488", "#ffffff", "#004488")) +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        strip.text = element_text(size = t.sizes[1]),
        strip.background = element_blank()) +
  xlab(bquote(italic(N[S]))) +
  ylab("Density")

ggsave("Figure3.pdf", f2, height = 10, width = 15)

```

# Figure 2: var vs difference between estimate and mean of simulations for mcmc and k; Figure S1
Note that the labels are flipped here, this produces figure 2

```{r Figure3}
# for all of the mpseg files, figure out if the percentage of points in the CI changes
## do for both each SNP and for Ns
ps_list <- ns_list <- vector("list", length(res_files))
for(i in 1:length(res_files)){
  tk <- res_files[[i]]$k
  titers <- res_files[[i]]$iters

  ps_prep <- prep_f1(pseg1 = pseg1[[paste0("K_", tk)]],
                     pseg2 = pseg2[[paste0("K_", tk)]], 
                     mpseg1 = res_files[[i]]$mpseg1,
                     mpseg2 = res_files[[i]]$mpseg2,
                     n1 = n1,
                     n2 = n2)

  ns_prep <- prep_f2(seg1 = res_files[[i]]$seg1, 
                     seg2 = res_files[[i]]$seg2, 
                     vseg1 = vseg1[[paste0("K_", tk)]],
                     vseg2 = vseg2[[paste0("K_", tk)]],
                     F1data = ps_prep,
                     n1 = n1,
                     n2 = n2)

  # per snp
  ps_prep$K <- tk
  ps_prep$iters <- titers
  ps_prep$sim_num <- i

  ## counts seg per snp
  cn1 <- res_files[[i]]$seg1
  cn1 <- rowSums(cn1)
  cn2 <- res_files[[i]]$seg2
  cn2 <- rowSums(cn2)

  cn <- rbind(data.frame(pop = paste0("N = ", n1/2), af = allele_frequencies, n_obs_seg = cn1),
              data.frame(pop = paste0("N = ", n2/2), af = allele_frequencies, n_obs_seg = cn2))
  ps_prep <- merge(ps_prep, cn, by = c("pop", "af"), sort = FALSE)

  # total
  ns_prep$all_mcmc_sums$K <- tk
  ns_prep$all_mcmc_sums$iters <- titers
  ns_prep$all_mcmc_sums$i <- i
  ns_prep$all_mcmc_sums$ENs <- ns_prep$sum_tab$nseg[match(ns_prep$all_mcmc_sums$pop,
                                                          ns_prep$sum_tab$pop)]

  ps_list[[i]] <- ps_prep
  ns_list[[i]] <- ns_prep$all_mcmc_sums
}

ps_all <- rbindlist(ps_list)

ps_all_m <- ps_all[source == "Mathematical",] # note: all of the CIs in this refer to: is the mathematical expectation in the observed CI.
ps_all_s <- ps_all[source == "Simulation",]
ps_all_merge <- merge(ps_all_m[,-c("source")], 
                      ps_all_s[,-c("source")], 
                      by = c("af", "pop", "snp", "sim_num", "lower", "upper", "K", "iters"), 
                      suffixes = c(".mathematical", ".simulated"))
ps_all_merge[,diff := p.mathematical - p.simulated]

# note: this code does the reverse: is the observed number of seg sites for a locus a set of sims within the PI expected based on the expected prob
ps_all_merge[,pi_lower := qbinom(.05, iters, p.mathematical)]
ps_all_merge[,pi_upper := qbinom(.95, iters, p.mathematical)]
ps_all_merge[,in_pi := n_obs_seg.simulated >= pi_lower & n_obs_seg.simulated <= pi_upper]

# Figure S1: gamma vs error rate
ps_all_aggregate_for_K <- ps_all_merge[, lapply(.SD, mean), .SDcols = "in_ci.mathematical", by = .(K, iters, pop)]
FS1 <- ggplot(ps_all_aggregate_for_K, aes(x = K, y = in_ci.mathematical*100)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(iters~pop) +
  theme_bw() +
  # khroma::scale_color_batlow(discrete = TRUE, reverse = TRUE) +
  # khroma::scale_fill_batlow(discrete = TRUE, reverse = TRUE) +
  theme(axis.text = element_text(size = t.sizes[2] - 2),
        axis.title = element_text(size = t.sizes[1] - 2),
        strip.text = element_text(size = t.sizes[1] - 2),
        strip.background = element_blank(),
        legend.title = element_text(size = t.sizes[1] - 2),
        legend.text = element_text(size = t.sizes[2] - 2)) +
  xlab(parse(text = "gamma")) +
  ylab("% estimates in CI")

save_plot("FigureS1.pdf", FS1, base_height = 15, base_width = 15)
save_plot("FigureS1.jpg", FS1, base_height = 15, base_width = 15)

# # Figure S2: maf vs error rate
# ps_all_aggregate_for_af <- ps_all_merge[, lapply(.SD, mean), .SDcols = "in_ci.mathematical", by = .(af, iters, pop)]
# FS2 <- ggplot(ps_all_aggregate_for_af, aes(x = af, y = in_ci.mathematical*100)) +
#   geom_point() +
#   geom_smooth(method = "lm", se = FALSE) +
#   facet_grid(iters~pop) +
#   theme_bw() +
#   # khroma::scale_color_batlow(discrete = TRUE, reverse = TRUE) +
#   # khroma::scale_fill_batlow(discrete = TRUE, reverse = TRUE) +
#   theme(axis.text = element_text(size = t.sizes[2]),
#         axis.title = element_text(size = t.sizes[1]),
#         strip.text = element_text(size = t.sizes[1]),
#         strip.background = element_blank(),
#         legend.title = element_text(size = t.sizes[1]),
#         legend.text = element_text(size = t.sizes[2])) +
#   xlab("Minor Allele Frequency") +
#   ylab("% estimates in CI")

ps_all_merge[,in_ci.mathematical := ifelse(in_ci.mathematical, "Yes", "No")]
ps_all_merge[,in_ci.mathematical := factor(in_ci.mathematical, c("Yes", "No"))]
f3a <- ggplot(ps_all_merge, aes(x = K, y = diff, color = in_ci.mathematical, group = in_ci.mathematical)) +
  geom_point(position = position_dodge(4)) +
  theme_bw() +
  facet_grid(~pop) +
  khroma::scale_color_highcontrast() +
  scale_x_continuous(breaks = k) +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        strip.text = element_text(size = t.sizes[1]),
        strip.background = element_blank(),
        legend.title = element_text(size = t.sizes[1]),
        legend.text = element_text(size = t.sizes[2])) +
  xlab(parse(text = "gamma")) +
  ylab(bquote(italic(hat(P)*"("*S[jq]*")") - italic(P*"("*S[jq]*")"))) +
  guides(color = guide_legend(title = "In 95% CI"))

f3b <- ggplot(mapping = aes(x = af, y = diff, color = in_ci.mathematical)) +
  geom_point(data = ps_all_merge[in_ci.mathematical == "Yes",]) +
  geom_point(data = ps_all_merge[in_ci.mathematical == "No",]) +
  theme_bw() +
  facet_grid(~pop) +
  khroma::scale_color_highcontrast() +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        strip.text = element_blank(),
        axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
        strip.background = element_blank(),
        legend.title = element_text(size = t.sizes[1]),
        legend.text = element_text(size = t.sizes[2])) +
  xlab("Minor Allele Frequency") +
  ylab(bquote(italic(hat(P)*"("*S[jq]*")") - italic(P*"("*S[jq]*")")))

f3c <- ggplot(ps_all_merge, aes(x = iters, y = diff, color = in_ci.mathematical, group = in_ci.mathematical)) +
  geom_point(position = position_dodge(400)) +
  theme_bw() +
  facet_grid(~pop) +
  khroma::scale_color_highcontrast() +
  scale_x_continuous(breaks = mcmcs) +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        strip.text = element_blank(),
        axis.text.x = element_text(size = t.sizes[2], 
                                   angle = 90, vjust = .5),
        strip.background = element_blank(),
        legend.title = element_text(size = t.sizes[1]),
        legend.text = element_text(size = t.sizes[2])) +
  xlab("Number of Simulated Draws") +
  ylab(bquote(italic(hat(P)*"("*S[jq]*")") - italic(P*"("*S[jq]*")")))

f3l <- ggpubr::get_legend(f3a)
f3 <- plot_grid(f3a + guides(color = "none"), 
          f3b + guides(color = "none"), 
          f3c + guides(color = "none"),
          ncol = 1, 
          align = "bl", 
          axis = "bl")

f3 <- plot_grid(f3, f3l, nrow = 1, rel_widths = c(1, .2))

save_plot("Figure2.pdf", f3, base_height = 11, base_width = 15)
```
# Figure 4 Nseg vs parameters; Figure S2

```{r Figure4}
ns_all <- rbindlist(ns_list)

ns_all$diff <- ns_all$ENs - ns_all$nseg

f4a <- ggplot(ns_all[iters == 10000,], aes(x = K, y = nseg)) +
  geom_errorbar(data = unique(ns_all[iters == 10000, c("pop", "K", "iters", "i", "lower", "upper")]),
            aes(x = K + 3.7, ymin = lower, ymax = upper), color = "#DDAA33", size = 1,
            inherit.aes = FALSE, width = 2) +
  geom_point(data = unique(ns_all[iters == 10000, c("pop", "K", "iters", "i", "lower", "upper", "ENs")]),
             aes(x = K + 3.7, y = ENs), color = "#DDAA33", size = 7, shape = 45,
            inherit.aes = FALSE) +
  stat_ydensity(aes(group = K), draw_quantiles = c(.025, .5, .975), scale = "width", bw = .4,
                fill = NA, width = 6) +
  facet_wrap(~pop) +
  scale_y_continuous(breaks = seq(40, 100, by = 10), limits = c(40, 100)) +
  theme_bw() + 
  scale_x_continuous(breaks = k) +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        legend.title = element_text(size = t.sizes[1]),
        strip.text = element_text(size = t.sizes[1]),
        strip.background = element_blank(),
        legend.text = element_text(size = t.sizes[2])) +
  ylab(bquote(italic(N[S]))) +
  xlab(bquote(gamma))

f4b <- ggplot(ns_all[K == 50,], aes(x = iters, y = nseg)) +
  geom_errorbar(data = unique(ns_all[K == 50,c("pop", "K", "iters", "i", "lower", "upper")]),
            aes(x = iters + 370, ymin = lower, ymax = upper), color = "#DDAA33", size = 1,
            inherit.aes = FALSE, width = 200) +
  geom_point(data = unique(ns_all[K == 50, c("pop", "K", "iters", "i", "lower", "upper", "ENs")]),
             aes(x = iters + 370, y = ENs), color = "#DDAA33", size = 7, shape = 45,
            inherit.aes = FALSE) +
  stat_ydensity(aes(group = iters), draw_quantiles = c(.025, .5, .975), scale = "width", bw = .4,
                fill = NA, width = 600) +
  facet_wrap(~pop) +
  theme_bw() + 
  scale_y_continuous(breaks = seq(40, 100, by = 5), limits = c(80, 100)) +
  scale_x_continuous(breaks = mcmcs) +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        legend.title = element_text(size = t.sizes[1]),
        axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
        strip.text = element_blank(),
        strip.background = element_blank(),
        legend.text = element_text(size = t.sizes[2])) +
  ylab(bquote(italic(N[S]))) +
  xlab("Number of Simulated Draws")

f4 <- plot_grid(f4a, f4b, ncol = 1, axis = "l", align = "l")
save_plot("Figure4.pdf", f4, base_height = 11, base_width = 15)

# S2
ns_all_agg <- ns_all[,lapply(.SD,mean), .SDcols = "nseg", by = c("ENs", "pop", "K", "iters", "i")]
levs <- paste0("Sims = ", sort(unique(ns_all_agg$iters)))
ns_all_agg$iters <- factor(paste0("Sims = ", ns_all_agg$iters), levels = levs)
ns_all_agg$pop <- gsub("n", "N", ns_all_agg$pop)
ns_all_agg$stat <- "Mean"

ns_all_agg_med <- ns_all[,lapply(.SD,median), .SDcols = "nseg", by = c("ENs", "pop", "K", "iters", "i")]
levs <- paste0("Sims = ", sort(unique(ns_all_agg_med$iters)))
ns_all_agg_med$iters <- factor(paste0("Sims = ", ns_all_agg_med$iters), levels = levs)
ns_all_agg_med$pop <- gsub("n", "N", ns_all_agg_med$pop)
ns_all_agg_med$stat <- "Median"

ns_all_agg <- rbind(ns_all_agg, ns_all_agg_med)
FS2 <- ggplot(ns_all_agg, aes(x = K, y = ENs - nseg, color = pop)) +
  geom_point() +
  facet_grid(stat~iters) +
  khroma::scale_color_highcontrast() +
  geom_hline(yintercept = 0) +
  theme_bw() +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        legend.title = element_text(size = t.sizes[1]),
        axis.text.x = element_text(size = t.sizes[2], angle = 90, vjust = .5),
        strip.background = element_blank(),
        strip.text = element_text(size = t.sizes[1]),
        legend.text = element_text(size = t.sizes[2])) +
  ylab(bquote(italic(E(N[S])) * " - " * italic(N[S,observed]))) +
  xlab(bquote(gamma)) +
  guides(color = guide_legend(title = "Population"))

save_plot("FigureS2.pdf", FS2, base_height = 10, base_width = 21)
save_plot("FigureS2.jpg", FS2, base_height = 10, base_width = 21)

```

```{r}
# show difference in HWE divergence with different combinations of 5 minor alleles in 100 genotypes
exd <- rbind(c(rep(0, 45), rep(1, 5)),
             c(rep(0, 46), 2, rep(1, 3)),
             c(rep(0, 47), rep(2, 2), 1))
exds <- import.snpR.data(as.data.frame(exd), snp.meta = data.frame(snp = 1:3), mDat = -9)
exds <- calc_hwe(exds)
get.snpR.stats(exds, stats = "hwe")

exc <- counts(exd)
rarefact(exc, 10)
```

#========real data example, mathematically============

```{r}
library(snpR)
set.seed(12321)
d <- readRDS("../data/monarchs_non_strict.RDS")
d <- d[sample(nrow(d), 1000),] # subsample 1k loci

# remove smaller than 10, although it works fine including stuff like SAI with only 4
sample.meta(d)$pop[sample.meta(d)$pop %in% c("ENA", "WNA")] <- "NAM"
ps <- summarize_facets(d, "pop")$pop
d <- d[pop = -c(names(ps)[which(ps < 10)])]

d <- calc_seg_sites(d, "pop", g = 0) # set gamma to the smallest N per loci
seg_res_mon <- get.snpR.stats(d, "pop", "seg_sites") # follows biogeographic history

# save this as a VCF for the supplementary material example
format_snps(d, "vcf", outfile = "../data/example_vcf.vcf", chr = "CHROM")
fwrite(sample.meta(d), "../data/example_metadata.txt", col.names = TRUE, row.names = FALSE, sep = "\t")
```

#=========real data example, via rarefaction==========
```{r}
# setup
sn <- format_snps(d, "sn", interpolate = FALSE)
sn <- sn[,-c(1:7)]
tmcmcs <- 100

pops <- substr(colnames(sn), 1, 3)
upops <- unique(pops)
mon_seg <- array(0, c(nrow(sn), tmcmcs, length(unique(pops))))

# do the rarefaction
for(i in 1:tmcmcs){
  cat("\nIter:", i)
  for(j in 1:nrow(sn)){
    # cat("\n\t", j)
    tg <- min(tapply(unlist(sn[j,]), pops, function(x) sum(!is.na(x)))) # set gamma for this locus
    for(k in 1:length(upops)){
      mon_seg[j,i,k] <- rarefact_mcmc(sn[j,which(pops == upops[k])], k = tg) # rarefact this locus in this pop
    }
  }
}

# summarize results
mon_rare_res <- matrix(0, tmcmcs, length(upops))
for(k in 1:length(upops)){
  mon_rare_res[,k] <- colSums(mon_seg[,,k])
}
colnames(mon_rare_res) <- upops
mon_rare_res <- as.data.table(mon_rare_res)
mon_rare_res$i <- 1:tmcmcs
mon_rare_res <- melt(mon_rare_res, id.vars = "i", variable.name = "pop", value = "Ns")

FS3 <- ggplot(mon_rare_res, aes(x = pop, y = Ns)) +
  geom_boxplot() +
  geom_point(alpha = .2, position = position_jitter(height = 0, width = .2)) +
  geom_point(data = seg_res_mon$weighted.means, aes(x = subfacet, y = seg_sites), color = "red", size = 4) +
  theme_bw() +
  ylab(bquote(italic(N[s]))) +
  xlab("Population") +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        legend.title = element_text(size = t.sizes[1]))

save_plot("FigureS3.pdf", FS3, base_width = 15, base_height = 11)
save_plot("FigureS3.jpg", FS3, base_width = 15, base_height = 11)
```