### 2024-09-30 scripts/ms_supp.RMD

---
title: "Seg_sites_rarefaction"
output: html_document
date: "2023-08-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE); library(snpR); library(ggplot2); library(data.table);
library(cowplot)
```

# Define parameters
```{r}
# define parameters
seed <- 231266
threads <- 25
k <- seq(10, 100, by = 10) # numbers to rarefact to
mcmcs <- seq(1000, 10000, by = 1000) # number of mcmc draws
max_missing <- 30 # maximum percentage of missing data in a locus, from a uniform dist
n1 <- 2500
n2 <- 250

## derived
set.seed(seed)

## plotting
colors <- khroma::color("highcontrast")(3)
ci_col <- as.character(colors[3])
colors <- as.character(colors[c(2,1)])
t.sizes <- c(20, 18)
```

# Setup genotypes
```{r}
# g <- GeneArchEst::process_ms("data/ms_out.ms", 10000000)
# 
# g <- GeneArchEst::convert_2_to_1_column(g$x)
# g <- t(g)
# 
# g1 <- g[,1:n1]
# g2 <- g[,(n1 + 1):(n1 + n2)]
# 
# # derived parms
# nal <- nrow(g1)
# miss_percent <- floor(runif(nal, 0, max_missing))/100
# 
# # add missing data
# ms1 <- lapply(n1*miss_percent, function(x) sample(n1, x, FALSE))  # which are missing
# ms2 <- lapply(n2*miss_percent, function(x) sample(n2, x, FALSE))
# 
# # populate missing
# 
# ## do for every locus
# for(i in 1:nal){ 
#   g1[i,ms1[[i]]] <- NA
#   g2[i,ms2[[i]]] <- NA
# }
# 
# # convert ns to gene copy numbers
# n1 <- n1*2
# n2 <- n2*2

```

# Rarefact mathematically
```{r}
set.seed(seed)

gd <- readRDS("../data/g_ms_large.RDS")
g1 <- gd$g1
g2 <- gd$g2

nal <- nrow(g1)

keep_snps <- sort(sample(nal, 10000, FALSE))
g1 <- g1[keep_snps,]
g2 <- g2[keep_snps,]

rm(gd); gc();

nal <- length(keep_snps)
allele_frequencies_g1 <- rowSums(g1, na.rm = TRUE)/(rowSums(!is.na(g1))*2)
allele_frequencies_g1[allele_frequencies_g1 > .5] <- 1 - allele_frequencies_g1[allele_frequencies_g1 > .5]

allele_frequencies_g2 <- rowSums(g2, na.rm = TRUE)/(rowSums(!is.na(g2))*2)
allele_frequencies_g2[allele_frequencies_g2 > .5] <- 1 - allele_frequencies_g2[allele_frequencies_g2 > .5]

counts <- function(g){
  return(data.frame(majhom = rowSums(g == 0, na.rm = TRUE), 
                    het = rowSums(g == 1, na.rm = TRUE), 
                    minhom = rowSums(g == 2, na.rm = TRUE)))
}
rarefact <- function(counts, k){
  if(length(k) > 1){
    warning("length(k) > 1, using first element only.\n")
    k <- k[1]
  }
  return(1 - ((choose(counts[,1], k) + choose(counts[,3], k))/choose(rowSums(counts), k)))
}
varseg <- function(p) p*(1-p)

c1 <- counts(g1) # geno counts
c2 <- counts(g2)

# rarefact
pseg1 <- lapply(k, function(tk) rarefact(c1, tk))
pseg2 <- lapply(k, function(tk) rarefact(c2, tk))
names(pseg1) <- names(pseg2) <- paste0("K_", k)

vseg1 <- lapply(pseg1, varseg)
vseg2 <- lapply(pseg2, varseg)

p1 <- lapply(pseg1, function(p) sum(p)/nal)
p2 <- lapply(pseg2, function(p) sum(p)/nal)
```

# Rarefact via simulations
Note: this chunk should run on a cluster or somewhere else allowing for large numbers of threads, then be re-imported in the next chunk. `q` should be passed as an argument for the iteration, as should `outfile`.

```{r cluster_MCMCs}
# set.seed(seed + q) # reset seed for cluster

rarefact_mcmc <- function(g, k){
  if(length(k) > 1){
    warning("length(k) > 1, using first element only.\n")
    k <- k[1]
  }

  miss <- is.na(g)
  if(any(miss)){
    g <- g[-which(miss)]
  }

  samp <- sample(g, k, FALSE)
  seg <- ifelse(any(samp == 1), 1, # got hets
                ifelse(any(samp == 0) & any(samp == 2), 1, 0)) # got both homs
  return(seg)
}

# # draw for each k/mcmc count combination
combs <- expand.grid(k, mcmcs)
colnames(combs) <- c("k", "iters")
# 
# tmcmcs <- combs$iters[q]
# tk <- combs$k[q]
# 
# seg1 <- seg2 <- matrix(0, nal, tmcmcs)
# for(i in 1:tmcmcs){
#   for(j in 1:nal){
#     seg1[j,i] <- rarefact_mcmc(g1[j,], k = tk)
#     seg2[j,i] <- rarefact_mcmc(g2[j,], k = tk)
#   }
# }
# 
# mpseg1 <- binom::binom.confint(rowSums(seg1), rep(tmcmcs, nrow(seg1)),
#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
# mpseg2 <- binom::binom.confint(rowSums(seg2), rep(tmcmcs, nrow(seg1)),
#                                method = "agresti-coull")[,c("mean", "lower", "upper")]
# 
# colnames(mpseg1) <- colnames(mpseg2) <- c("mean", "lower", "upper")
# 
# res <- list(seg1 = seg1, seg2 = seg2, mpseg1 = mpseg1, mpseg2 = mpseg2, k = tk, iters = tmcmcs)
# 
# saveRDS(res, paste0(outfile, "_", q, ".RDS"))

```

# import and condense simulation results
```{r}
res_files <- list.files("../data/ms_MCMC/", "mcmc.+\\.RDS", full.names = TRUE)
q <- gsub(".+__", "", res_files)
q <- as.numeric(gsub("\\.RDS", "", q))
# res_files <- pbapply::pblapply(res_files, readRDS)
ks <- combs$k[q]
n_iters <- combs$iters[q]
# ks <- unlist(purrr::map(res_files, "k"))
# n_iters <- unlist(purrr::map(res_files, "iters"))

```

# Figure 1: Pseg for four individual runs (k = 10 and 100, N = 100, N = 1000)

```{r Figure1}
n1 <- n1*2
n2 <- n2*2
prep_f1 <- function(pseg1, pseg2, mpseg1, mpseg2, n1, n2){
  #========plot the empirical CIs around every loci vs the predicted value==============
  # plotting the empirical CIs vs the estimated mean
  pseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies_g1),
                data.frame(pop = paste0("N = ", n2/2), p = pseg2, snp = 1:nal, af = allele_frequencies_g2))

  oseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = mpseg1[,1], snp = 1:nal, af = allele_frequencies_g1),
                data.frame(pop = paste0("N = ", n2/2), p = mpseg2[,1], snp = 1:nal, af = allele_frequencies_g2))

  pseg$source <- "Mathematical"
  oseg$source <- "Simulation"

  seg <- rbind(pseg, oseg)
  cis <- rbind(cbind.data.frame(mpseg1, pop = paste0("N = ", n1/2), af = allele_frequencies_g1, snp = 1:nal),
               cbind.data.frame(mpseg2, pop = paste0("N = ", n2/2), af = allele_frequencies_g2, snp = 1:nal))
  colnames(cis) <- c("p", "lower", "upper", "pop", "af", "snp")
  seg <- merge(seg, cis[,c("lower", "upper", "pop", "snp")], by = c("snp", "pop"))
  seg$in_ci <- ifelse(seg$p >= seg$lower & seg$p <= seg$upper, TRUE, FALSE)

  return(seg)
}

K10_data <- which(ks == 10 & n_iters == 10000)
K10_data <- readRDS(res_files[[K10_data]])
F1data_K10 <- prep_f1(pseg1 = pseg1$K_10, 
                      pseg2 = pseg2$K_10, 
                      mpseg1 = K10_data$mpseg1,
                      mpseg2 = K10_data$mpseg2,
                      n1 = n1,
                      n2 = n2)

K100_data <- which(ks == 100 & n_iters == 10000)
K100_data <- readRDS(res_files[[K100_data]])

F1data_K100 <- prep_f1(pseg1 = pseg1$K_100, 
                       pseg2 = pseg2$K_100, 
                       mpseg1 = K100_data$mpseg1,
                       mpseg2 = K100_data$mpseg2,
                       n1 = n1,
                       n2 = n2)

F1data_K10$K <- "gamma * ' = 10'"
F1data_K100$K <- "gamma * ' = 100'"

F1data <- rbind(F1data_K10, F1data_K100)
F1data$pop <- paste0("'", F1data$pop, "'")

F1data$in_ci <- ifelse(F1data$in_ci, "Yes", "No")

f1 <- ggplot(F1data[F1data$source == "Mathematical",],
             aes(x = af, y = p)) +
  geom_point(aes(color = in_ci), size = 2.5) +
  geom_errorbar(aes(ymax = upper, ymin = lower)) +
  facet_grid(K~pop, labeller = label_parsed) +
  theme_bw() +
  theme(strip.background = element_blank()) +
  scale_color_manual(values = colors) +
  guides(color = guide_legend(title = "In 95% CI")) +
  ylab(bquote(italic(hat(P)*"("*S[jq]*")"))) +
  xlab("Minor Allele Frequency") +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        strip.text = element_text(size = t.sizes[1]),
        legend.title = element_text(size = t.sizes[1]),
        legend.text = element_text(size = t.sizes[2]))

ggsave("FigureS4.pdf", f1, width = 15, height = 10)
tapply(F1data[F1data$source == "Mathematical", "in_ci"], F1data[F1data$source == "Mathematical", c("K", "pop")], mean) # percent in interval
```

# Figure 3: Distribution of Nseg vs Eseg for four runs
Note that the label are flipped here: this produces figure 3.

```{r}
prep_f2 <- function(seg1, seg2, vseg1, vseg2, F1data, n1, n2){
  seg <- F1data

  # normal approximation CI and PI equation given var and N
  nap_pi <- function(alpha, vseg, n) qnorm(1 - (alpha/2))*sqrt(vseg)*sqrt(1 + (1/n))
  nap_ci <- function(alpha, vseg, n) qnorm(1 - (alpha/2))*sqrt(vseg)/sqrt(n)

  # sums across loci
  msum1 <- colSums(seg1)
  msum2 <- colSums(seg2)

  # CI/PI
  msum_ci1 <- nap_ci(.05, var(msum1), length(msum1))
  msum_ci2 <- nap_ci(.05, var(msum2), length(msum2))

  sum_pi1 <- nap_pi(0.05, sum(vseg1), nal)
  sum_pi2 <- nap_pi(0.05, sum(vseg2), nal)

  # add to table
  msum_tab <- data.frame(nseg = c(mean(msum1), mean(msum2)),
                         lower = c(mean(msum1) - msum_ci1, mean(msum2) - msum_ci2),
                         upper = c(mean(msum1) + msum_ci1, mean(msum2) + msum_ci2),
                         pop = c(paste0("n = ", n1/2), paste0("n = ", n2/2)))

  sum_tab <- tapply(seg[seg$source == "Mathematical",]$p, seg[seg$source == "Mathematical",]$pop, sum)
  sum_tab <- as.data.frame(sum_tab)
  sum_tab$pop <- tolower(rownames(sum_tab))
  colnames(sum_tab)[1] <- "nseg"
  sum_tab <- merge(sum_tab, msum_tab[,c("pop", "lower", "upper")], by = "pop")
  sum_tab$in_ci <- sum_tab$nseg <= sum_tab$upper & sum_tab$nseg >= sum_tab$lower

  # add pi
  all_mcmc_sums <- rbind(cbind.data.frame(nseg = msum1, pop = paste0("n = ", n1/2)),
                         cbind.data.frame(nseg = msum2, pop = paste0("n = ", n2/2)))
  pi_tab <- data.frame(pop = c(paste0("n = ", n1/2), paste0("n = ", n2/2)),
                       lower = c(sum_tab$nseg[sum_tab$pop == paste0("n = ", n1/2)] - sum_pi1,
                                 sum_tab$nseg[sum_tab$pop == paste0("n = ", n2/2)] - sum_pi2),
                       upper = c(sum_tab$nseg[sum_tab$pop == paste0("n = ", n1/2)] + sum_pi1,
                                 sum_tab$nseg[sum_tab$pop == paste0("n = ", n2/2)] + sum_pi2))
  all_mcmc_sums <- merge(all_mcmc_sums, pi_tab, by = "pop")
  all_mcmc_sums$in_pi <- all_mcmc_sums$nseg <=  all_mcmc_sums$upper & all_mcmc_sums$nseg >=  all_mcmc_sums$lower

  return(list(sum_tab = sum_tab, all_mcmc_sums = all_mcmc_sums, pi_tab = pi_tab))
}

K20_data <- which(ks == 20 & n_iters == 10000)
K20_data <- readRDS(res_files[[K20_data]])

F1data_K20 <- prep_f1(pseg1 = pseg1$K_20, 
                       pseg2 = pseg2$K_20, 
                       mpseg1 = K20_data$mpseg1,
                       mpseg2 = K20_data$mpseg2,
                       n1 = n1,
                       n2 = n2)

F2data_K10 <- prep_f2(seg1 = K10_data$seg1, 
                      seg2 = K10_data$seg2, 
                      vseg1 = vseg1$K_10,
                      vseg2 = vseg2$K_10,
                      F1data = F1data_K10,
                      n1 = n1,
                      n2 = n2)

F2data_K20 <- prep_f2(seg1 = K20_data$seg1, 
                       seg2 = K20_data$seg2, 
                       vseg1 = vseg1$K_20,
                       vseg2 = vseg2$K_20,
                       F1data = F1data_K20,
                       n1 = n1,
                       n2 = n2)

F2data_K10$sum_tab$K <- "gamma * ' = 10'"
F2data_K20$sum_tab$K <- "gamma * ' = 20'"
F2data_K10$all_mcmc_sums$K <- "gamma * ' = 10'"
F2data_K20$all_mcmc_sums$K <- "gamma * ' = 20'"
F2data_K10$pi_tab$K <- "gamma * ' = 10'"
F2data_K20$pi_tab$K <- "gamma * ' = 20'"

F2data_sums <- rbind(F2data_K10$sum_tab, F2data_K20$sum_tab)
F2data_sums$pop <- gsub("n", "N", F2data_sums$pop)
F2data_sums$pop <- paste0("'", F2data_sums$pop, "'")

F2data_mcmcs <- rbind(F2data_K10$all_mcmc_sums, F2data_K20$all_mcmc_sums)
F2data_mcmcs$pop <- gsub("n", "N", F2data_mcmcs$pop)
F2data_mcmcs$pop <- paste0("'", F2data_mcmcs$pop, "'")

F2data_pi <- rbind(F2data_K10$pi_tab, F2data_K20$pi_tab)
F2data_pi$pop <- gsub("n", "N", F2data_pi$pop)
F2data_pi$pop <- paste0("'", F2data_pi$pop, "'")

F2data_mcmcs <- as.data.table(F2data_mcmcs)
levs <- unique(F2data_mcmcs[,c("pop", "K")])
F2data_densities <- vector("list", nrow(levs))

for(i in 1:nrow(levs)){
  d <- density(F2data_mcmcs[K == levs$K[i] & pop == levs$pop[i],]$nseg)
  q <- quantile(F2data_mcmcs[K == levs$K[i] & pop == levs$pop[i],]$nseg, c(0.025,0.975))
  F2data_densities[[i]] <- data.table(x = d$x, y = d$y, K = levs$K[i], pop = levs$pop[i])
  F2data_densities[[i]]$quant <- factor(findInterval(F2data_densities[[i]]$x, q))
}

F2data_densities <- rbindlist(F2data_densities)

f2 <- ggplot(F2data_densities, aes(x = x, y = y)) +
  geom_line() +
  geom_ribbon(aes(ymin = 0, ymax = y, fill = quant), show.legend = FALSE) +
  geom_vline(data = F2data_sums, aes(xintercept = nseg), color = colors[1]) +
  geom_vline(data = F2data_pi, aes(xintercept = lower), linetype = "dashed", color = colors[2]) +
  geom_vline(data = F2data_pi, aes(xintercept = upper), linetype = "dashed", color = colors[2]) +
  facet_grid(K~pop, labeller = label_parsed) +
  theme_bw() +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = c("#004488", "#ffffff", "#004488")) +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        strip.text = element_text(size = t.sizes[1]),
        strip.background = element_blank()) +
  xlab(bquote(italic(S))) +
  ylab("Density")

ggsave("FigureS5.svg", f2, height = 10, width = 15)

```